---
title: "ALB-IO test SAMtools SP_SS"

author: 
 - Henning Winker (GFCM) 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    number_sections: yes
    toc: yes
    #keep_tex: false
vignette: >
  %\VignetteIndexEntry{ss3daigs Vignetted}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r wrap-hook, echo = FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```


```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "  " ,fig.align = 'center', cache=FALSE,tidy.opts=list(width.cutoff=80), tidy=TRUE)

```




# Load packages


```{r, message=FALSE}
# Load
library(ggplot2)
library(openMSE)
library(FishLife)
library(SPMpriors)
library(FLCore)
library(ggplotFL)
library(FLRef)
library(FLife)
library(JABBA)
library(ggpubr)
```

# Life history parameters

Life history parameters of ALB-IO for females were assumed assumed as follows. 

```{r}
alb = flmvn_traits(Genus="Thunnus",Species="alalunga",h=c(0.6,0.9),Plot=F)$traits
```




```{r}
linf = alb[alb$trait=="Loo","mu.stk"] 
k =  alb[alb$trait=="K","mu.stk"] 
t0 = -0.5 
tm = alb[alb$trait=="tm","mu.stk"] 
a = 0.00001
b = 3.04
tmax =  ceiling(alb[alb$trait=="tmax","mu.stk"])
s = alb[alb$trait=="h","mu.stk"]


# check vbgf
la = data.frame(age=0:tmax,length=vonbert(linf,k,t0,age=0:tmax))
la

```

An instantaneous rate of natural mortality was estimated using Then based on the maximum age estimates `tmax1 = 22` and `tmax2 = 12`

```{r}
M = 4.899*tmax^-.916 
```


# Leslie matrix with stock-specitic direct input

The recent version of the `JABBA` package provides a Leslie matrix function for use stock specific estimates to get an expected value for $r$.

The first step is build a `jbio` object and then pass this object to compute $r$ (and Generation time) with `jbleslie()`

```{r}

bio = jbio(amax=tmax,nsexes=1,Loo=linf,k=k,t0=t0,aW=1000*a,bW=b,
     mat=c(tm,tm*1.2,1),M=M,h=s)
r = jbleslie(bio)$r
r

```

# Setting up a generic OM for ALB

The `FLife` function `lhPar()` is populated with some of basic life history estimates for growth,length-weight, maturity based on regional information and steepness from FishLife/SPMpriors.

```{r}

par=lhPar(FLPar(linf=linf,k=k,t0=t0,
            a=a,b=b,a50 =2,
            s=s,m1=M))
```

The `FLife` function `lhEql` is then used as a constructer for a basic `FLBRP()` object.

```{r}

eql=lhEql(par,range = c(min = 0, max = tmax, minfbar = 1, maxfbar = tmax-1, plusgroup=tmax))
```

Now further manipulations can be made, such as adjusting the inbuilt Gislason $M$ to a scaled Lorenzen $M$ and specifying a new maturity ogive. 


```{r}
# M
m(eql)[] = Mlorenzen(stock.wt(eql),Mref=M,Aref=2)

mat50 = tm
mat95 = tm*1.5
mat(eql) = newselex(mat(eql),FLPar(S50=mat50,mat95=mat95,Smax=1000,Dcv=0.5,Dmin=0.1))
```

It is important to then update the stock recruitment the stock recruitment relationship for a given steepness $s$ value if key biological parameters are modified.

```{r}
# check srr parameters
params(eql)
# update SRR given s after modifying m
eql = updsr(eql,s=s) 
params(eql)

```

After making all the desired modifications the time horizon can be adjusted with `fbar` range. Here a time horizon from 195ww1-2020 is chosen.

```{r}
# change time horizon with a fixed intial f0
f0 = 0.01
fbar(eql) = FLQuant(f0,dimnames=list(year=1951:2020))

```

Now `FLBRP` can be directly converted into an `FLStock` and the stock recruitment relationship can be extracted.


```{r}
stk = as(eql,"FLStock")
units(stk) = standardUnits(stk) # assign units
sr = as(as(eql,"predictModel"),"FLSR")
```

Once the stock is constructed further manipulations are straight forward, such as specifying fishing selectivity using the 5-parameter selectivity function `newselex`

A new selectivity can be set by adjusting the `harvest` slot. If some dome-shaping is desired, $Smax$ has to be smaller then the plus group, $Dcv$ governs the slope of the half-normal and $Dmin$ governs the hight at the right end tail.

```{r}

# catch before maturity
S50 = 3 
S95 = 4.2
# some Doming
Smax = 7
Dcv = 0.5
Dmin = 0.2

harvest(stk) =  f0*newselex(
  harvest(stk),FLPar(S50=S50,S95=S95,
    Smax=Smax,Dcv=Dcv,Dmin=Dmin))

```

The so generated age-structured relationships for weight-at-age, maturity-at-age, natural mortality and selective are shown in Fig. 1.

```{r fig1, fig.height=4,fig.width=5, fig.cap = paste0("Specified biological and selectivtiy function for natural mortality, selectivity, maturity and somatic growth in weight-at-age")}

ggplot(FLQuants(stk,"m","catch.sel","mat","catch.wt"))+
  geom_line(aes(age,data))+
  facet_wrap(~qname,scale="free")+theme_bw()

```

## Estimating priors from FLStock and SRR

First, it is straight forward to compute the intrinsic rate of population increase from a Leslie for a specified steepness value. However, this $r$ estimate ignores selectivity and should only be used in the context of a Schaefer model with $F_{MSY} = r/2$.  

```{r}
r.leslie = mean(productivity(stk,s=s)$r)
r.leslie
# compare
r
# Fmsy
r.leslie/2 
```

An alternative is to estimate $r$ and the shape parameter $n$ as function $MSY$, $VB_{MSY}$ and $VB_0$ from an age-structured equilibrium model (ASEM), where $BB$ denotes that the vulnarable (or exploitable) biomass as function of selectivity. The basic relationships for $r$ and $n$ can then be approximated as follows:

$$F_{MSY} = \frac{MSY}{VB_{MSY}}$$
$$\frac{VB_{MSY}}{EB_0} = n^{(-\frac{1}{n-1})}$$

$$r =F_{MSY}\frac{n-1}{1-n^{-1}}$$

To compute the expected values for $r$ and $n$ (here denoted as $m$), requires the stock biology and selectivity together with the SRR as input.

```{r}
brp = FLBRP(stk,sr)
r.pella = asem2spm(brp)

r.pella

```

This is very close to a Fox model with m = 1


```{r fig2, fig.height=5,fig.width=10, fig.cap = paste0("Specified biological and selectivtiy function for natural mortality, selectivity, maturity and somatic growth in weight-at-age")}

ggpubr::ggarrange(
plotpf(brp,rel=F),
plotpf(brp,rel=T),common.legend = T
)


```


## Simulating stock dynamics with evolutionary F-trajectories

The operating model is assumed to represent the "true" age-structured stock dynamics for evaluating the estimation accuracy of Spict. The "true" MSY based reference points can be easily added to the `FLStock` by extending it to `FLStockR`.

```{r}
# Estimate refpts
brp = computeFbrp(stk,sr,proxy="msy",blim=0.3,type="btgt")
stk = FLStockR(stk)
stk@refpts = Fbrp(brp)
# check
stk@refpts
```
```{r fig3, fig.height=7,fig.width=9, fig.cap = paste0("Equilibrium curves Recruitment, SSB, F and Landings and estimated MSY-based reference points for the ALB OM")}

ploteq(brp)

```

`FLRef` provides `rffwd` as an extension of the `FLCore` F-projection function`ffwd`, which enables to setup a control object for generating randomized evolutionary F-trajectories (Thorson et al. 2016). The parameters required are: (1) $Feq$ which sets the terminal depletion level baseline relative to $SB_0$, `Frate` which is determines the turn-over time of the `F-evalution`, `Fsigma` determines the annual variation in F and `SB0` is reference biomass level that is typically specified as $SSB_0$.

Here a one-way downhill is approximated

```{r}
its = 100 # iterations
b0 = an(refpts(eql)["virgin","ssb"]) # SSB0 for brp

control = FLPar(Feq=0.15,Frate=0.05,Fsigma=0.15,SB0=b0,
                minyear=dims(stk)$minyear+1,
                maxyear=dims(stk)$maxyear,its=its)
```

The stock dynamics are forecasted with random recruitment under and evolving F-trajectory.

```{r}
set.seed(1234)
# Random recruitment deviations with sigR = 0.5 and AR1 rho = 0.3
rec_devs = ar1rlnorm(0.3, 1951:2020, its, 0, 0.5)
# propagate desired iterations
stki <- propagate(stk, its)
# create OM 
om <- rffwd(stki,sr,
             control=control,
             deviances=rec_devs)
om@refpts = stk@refpts
```




```{r fig5, fig.height=7,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Simulated population dynamics compared to reference points from the off-the-shelf OM for ALB based on 100 iterations")}

plotAdvice(om)

```

The last step is to generate the observed survey index using the `FLRef` function `bioidx.sim()`. By default, the index assumes the same fishing selectivity in the stock object (but see `?bioidx.sim`. In this case, a moderate observation error of `sigma=0.25` is assumed and the catchability coefficient $q$ that scales the exploitable biomass to the index is set as `q = 0.001`. The index is assumed to start in 1999.  

The index is from LL CPUE an immolates the adult biomass without  doming

```{r}
cpue.sel = newselex(harvest(stk),FLPar(S50=3.5,mat95=4.2,
Smax=100,Dcv=Dcv,Dmin=0.05))
idx = window(bioidx.sim(om,sel=cpue.sel,sigma=0.25,q=0.001),start=1970)
```

```{r fig6, fig.height=7,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Individual iterations of simulated trajectories of $B/B_{MSY}$, $F/F_{MSY}$, $Catch/MSY$ and the normalized CPUE indices from the off-the-shelf OM  shown for the first 20 iterations")}

flqs= FLQuants("B/Bmsy"=ssb(om)/om@refpts["Btgt"],
               "F/Fmsy"=fbar(om)/om@refpts["Fmsy"],
               "Catch/MSY"=catch(om)/om@refpts["Yeq"]
               ,Index=idx@index%/%yearMeans(idx@index))

ggplot(iter(flqs,1:20))+
  geom_line(aes(year,data,col=ac(iter)),alpha=0.5)+
  theme_bw()+facet_wrap(~qname)+
  theme(legend.position = "none")+geom_hline(yintercept = 1,linetype=2)+
  scale_color_manual(values=c(rainbow(20)))

```

# Simulation testing with Spict (need to convert to SP_SS)

# SP_SS estimation scenarios


```{r}
#' jb2FLStockR()
#' @param jabba fit from JABBA fit_jabba() or jabba$kbtrj 
#' @param bfrac biomass limit reference point 
#' @param rel if TRUE ratios BBmsy and FFmsy are stored
#' @return FLStockR with refpts
#' @export
spss2FLStockR <- function(spss,bfrac=0.3,rel=FALSE){
  df = spss@TMB_report
  
  year = an(names(df$dynamic_SSB0))
  n = length(year)-1
  bdev = c(1,exp(df$log_B_dev))[1:n]
  year = year[1:n]
  N = as.FLQuant(data.frame(age=1,year=year,unit="unique",
            season="all",area="unique",iter=1,data=bdev))
  C = as.FLQuant(data.frame(age=1,year=year,unit="unique",
                            season="all",area="unique",iter=1,data=df$Cpred))
  
  if(!rel){
    Mat = as.FLQuant(data.frame(age=1,year=year,unit="unique",
        season="all",area="unique",iter=1,data=df$B[1:n]/bdev))
    
    H = as.FLQuant(data.frame(age=1,year=year,unit="unique",
                              season="all",area="unique",iter=1,data=df$F))
    B= as.FLQuant(data.frame(age=1,year=year,unit="unique",
               season="all",area="unique",iter=1,data=df$B[1:n]))
  } else {
    Mat = as.FLQuant(data.frame(age=1,year=year,unit="unique",
    season="all",area="unique",iter=1,data=(df$B[1:n]/df$BMSY)/bdev))
    
    H = as.FLQuant(data.frame(age=1,year=year,unit="unique",
              season="all",area="unique",iter=1,data=df$F/df$FMSY))
    B= as.FLQuant(data.frame(age=1,year=year,unit="unique",
      season="all",area="unique",iter=1,data=df$B[1:n]/df$BMSY))
    }
  
  
  stk = FLStockR(
    stock.n=N,
    catch.n = C,
    landings.n = C,
    discards.n = FLQuant(0, dimnames=list(age="1", year = (year))),
    stock.wt=FLQuant(1, dimnames=list(age="1", year = (year))),
    landings.wt=FLQuant(1, dimnames=list(age="1", year = year)),
    discards.wt=FLQuant(1, dimnames=list(age="1", year = year)),
    catch.wt=FLQuant(1, dimnames=list(age="1", year = year)),
    mat=Mat,
    m=FLQuant(0.0001, dimnames=list(age="1", year = year)),
    harvest = H,
    m.spwn = FLQuant(0, dimnames=list(age="1", year = year)),
    harvest.spwn = FLQuant(0.0, dimnames=list(age="1", year = year))
  )
  units(stk) = standardUnits(stk)
  stk@catch = computeCatch(stk)
  stk@landings = computeLandings(stk)
  stk@discards = computeStock(stk)
  stk@stock = as.FLQuant(data.frame(age=1,year=year,unit="unique",
                    season="all",area="unique",iter=1,data=df$B[1:n]))
  
    stk@refpts = FLPar(
      Fmsy = df$FMSY,
      Bmsy = df$BMSY,
      MSY = df$MSY,
      Blim= df$BMSY*bfrac,
      B0 = df$K)
  
  if(rel){
    stk@refpts[1:2] = 1
    stk@refpts["Blim"] = bfrac
    stk@refpts["B0"] = df$K/df$BMSY
  }
  
  return(stk)
}
```

The two scenarios explored were:

(1) `Schaefer` with r prior from Leslie matrix
(2) `Pella` with priors from ASPM (close to Fox)


Below a simple wrapper function is presented that: 

+ Fits the 2 Spict scenarios
+ Applies `spict2FLStockR` to convert Spict fits into simplified `FLStock` objects
+ Apply `stock2ratios` to convert OM into a simplified `FLStock` object
+ Combines all into a `FLStocks` object output


```{r}

runi <- function(om,idx,it){

# index
dfi =as.data.frame(iter(idx@index,it))
# catch 
dfc = as.data.frame(iter(om@catch,it))
# year
year = dfc$year
n = length(year)
inp = new("Data",y=year)
inp@Year = year
inp@Cat= inp@Ind =inp@CV_Ind= matrix(NA,1,n) 
inp@Cat[1,] = dfc$data 
inp@Ind[1,which(year%in%dfi$year)] = dfi$data 
inp@CV_Ind[] = 0.25

# Parameter init control
# MSY = 60
# Fmsy = 0.2
# R prior does not seem to work with sd =  CV*mu? or variance???
# Must be super precise for an imprecise estimate
fits=list(
Schaefer = SP_SS(Data = inp,fix_sigma = T ,use_r_prior = T,early_dev = c("all"),
                  start = list(MSY=50,dep = 0.95 , sigma = 0.2, 
                               tau = 0.05,n=2,r_prior=c(r.leslie,0.01))),
Pella = SP_SS(Data = inp,fix_sigma = T ,use_r_prior = T,early_dev = c("all"), 
            start = list(MSY=50,dep = 0.95, sigma = 0.2, tau = 0.05,
                         n=r.pella[[2]],r_prior=c(r.pella[[1]],0.01)))
)

res = FLStocks(lapply(fits,function(x){
 spss2FLStockR(x,rel=T) 
})) 
# cut forecast
flom = stock2ratios(iter(om,it))
res = FLStocks(c(om=flom,res))   
return(res)
}
```


```{r fig7, fig.height=10,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("OM and estimated trajectories of $B/B_{MSY}$ and $F/F_{MSY}$ for the first iteration")}

res = runi(om,idx,it=2)

plot(res,metrics=list("B/Bmsy"=ssb,"F/Fmsy"=fbar,Catch=landings))+
  geom_hline(yintercept = 1,linetype=2)+theme_bw()


```

\pagebreak

## Quick simulation performance comparison


```{r,echo=FALSE}
load(file="sim.alb.spss.rdata")

```


```{r,eval=FALSE}
sims = runi(om,idx,it=1)
sims =FLStocks(lapply(sims,function(x)propagate(x,its)))

for(i in 2:its){
  out=runi(om,idx,it=i)
  for(j in 1:length(out)) iter(sims[[j]],i) = out[[j]]} 

sims= FLStocks(Map(function(x,y){
  x@name=y
  x},x=sims,y=as.list(names(sims))))
```
```{r fig8, fig.height=7,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("OM and estimated trajectories of $B/B_{MSY}$ and $F/F_{MSY}$ for 100 iterations")}

plot(sims,metrics=list("B/Bmsy"=ssb,"F/Fmsy"=fbar))+
  geom_hline(yintercept = 1,linetype=2)+theme_bw()

```

Several runs failed to converge the hessian, but these failed runs were not quantified removed for this initial evaluation.

By making use of `FLQuants` in is reasonably straight forward compute some estimation error with respect to the "true" value of $B_{MSY}$ and $F_{MSY}$ for the terminal year 2020.  



```{r}
simt= window(sims,start=2015)
bias.b = FLQuants(lapply(simt[2:3],function(x){
  log(ssb(x))-log(window(ssb(simt$om)))
}))
bias.f = FLQuants(lapply(simt[2:3],function(x){
  log(fbar(x))-log(window(fbar(simt$om)))
}))

out = rbind(
  data.frame(as.data.frame(bias.b),what="BBmsy"),
  data.frame(as.data.frame(bias.f),what="FFmsy")
)
```



```{r fig9, fig.height=4,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Boxplot showing estimation errors of $log(B/B_{MSY})$ and $log(F/F_{MSY})$ for 2020 with respect to the true quantities of the OM")}

ggplot(out,aes(x=qname,data,fill=qname))+
  geom_boxplot(outlier.shape = NA)+
  facet_wrap(~what)+
  geom_hline(yintercept = 0,col=2)+theme_bw()+
  xlab("Scenario")+ylab("Error")+ylim(-2,2)

```

\pagebreak

```{r fig10a, fig.height=5,fig.width=8, fig.cap = paste0("Median Trajectories of $VB/VB_{MSY}$ and $SB/SB_{MSY} for 100 OM iteratsions")}

asem.vb = asem2spm(brp,quant="vb")

df1 = data.frame(as.data.frame(iterMedians(vb(om))/asem.vb["Bmsy"]),qname="vb")
df2 = data.frame(as.data.frame(iterMedians(ssb(om))/stk@refpts[2]),qname="ssb")
df = rbind(df1,df2)

ggplot(df,aes(year,data,color=qname))+geom_line()+
  theme_bw()+ylab(expression(B/B[MSY]))+geom_hline(yintercept = 1,linetype=2)+
  theme(legend.title = element_blank())

```
